---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# SimpleFinCalc

<!-- badges: start -->
<!-- badges: end -->

The 'SimpleFinCalc' R package provides a straightforward suite of functions for performing common financial calculations, making it an excellent educational tool or quick-reference utility. It includes core functions for analyzing the time value of money, such as computing future and present values, as well as calculating loan payments. The package also features a function for generating a full amortization schedule, allowing users to easily break down loan payments into principal and interest over time. By encapsulating complex financial formulas into easy-to-use functions, SimpleFinCalc helps users perform essential financial analyses without the need for manual calculations, streamlining tasks for students, analysts, or anyone interested in personal finance.

The 'SimpleFinCalc' R package includes a built in data set 'profiles1' to assist in the testing and demonstration of the many functions that the packages includes. The data set is available when the package is installed and the library is loaded in to the R working environment. The data set can alo be referenced with the data() function.

## Installation

You can install the development version of SimpleFinCalc from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("rwalshUSF/SimpleFinCalc")
```

# Data
Here is a summary of the variables in the built-in 'profiles1' data set:

```{r profiles1}
library(SimpleFinCalc)
summary(profiles1)
```

# Examples of the package's functions with user input and 'profiles1' data

## --- 1. Time Value of Money (TVM) Functions ---

### Calculate the Future Value (FV) of a lump-sum investment (simple interest).

Simple Future Value - $FV = PV * (1 + r)^n$

```{r}
library(SimpleFinCalc)
# The 'fv_simple()' function:
#
# This function performs a 'Simple Future Value' calculation with robust input 
# validation. It takes three variables; Principal, Rate, and Years, and returns 
# the calculated value.

# Using user variables:
testP<-1000
testR<-5
testY<-10
fv_simple(testP,testR,testY)
# Therefore if a person had a Principal of $1000, at a Rate of .05% for 10 Years 
# they would have ~$1628.895 in the future!

# Using a record from the 'profiles1' data set

# Print the first record in the data set
profiles1[1,]

# Call the function using the 'profiles1[1]' data
fv_simple(profiles1$Principal[1],profiles1$Rate[1],profiles1$Years[1])
# Therefore if a person had a Principal of $457488.2, at a Rate of 0.13506535% 
# for 25 Years they would have ~$10861782.00 in the future!
```
