---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# SimpleFinCalc

<!-- badges: start -->

<!-- badges: end -->

The goal of SimpleFinCalc is to provide a straightforward suite of functions for performing common financial calculations, making it an excellent educational tool or quick-reference utility. It includes core functions for analyzing the time value of money, such as computing future and present values, as well as calculating loan payments. The package also features a function for generating a full amortization schedule, allowing users to easily break down loan payments into principal and interest over time. By encapsulating complex financial formulas into easy-to-use functions, SimpleFinCalc helps users perform essential financial analyses without the need for manual calculations, streamlining tasks for students, analysts, or anyone interested in personal finance.

## Vignette
```{r child-chunk, child = "vignettes/SimpleFinCalc.Rmd"}

```
